Clines from PDE

### 2015-10-25 notes/pde-clines.Rmd
allowed logA in tran.1D

Here's a function to solve this numerically:
```{r solve_pde_setup, cache=TRUE}
thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 5000)
solve.pde <- function ( u, v=u, r, times, grid, ... ) {
    vfn <- function(x,t){v(x=x,t=t,...)}
    cline.A <- setup.prop.1D(grid=grid, func=u, ...)
    cline.B <- setup.prop.1D(grid=grid, func=function(x,...)1-u(x,...), ...)
    yinit <- c( rep(1.0,grid$N), rep(0.0,grid$N) )
    r <- 0.1
    pde.fn <- function (t,y,parms,...) {
        yA <- y[1:grid$N]
        yB <- y[grid$N+(1:grid$N)]
        tran.A <- tran.1D(C=yA, A=cline.A, D=1/2, dx=grid)$dC
        tran.B <- tran.1D(C=yB, A=cline.B, D=1/2, dx=grid)$dC
        list( c( 
                tran.A + r * (1-vfn(x=grid$x.mid,t=t)) * (yB-yA),
                tran.B + r * vfn(x=grid$x.mid,t=t) * (yA-yB)
                ) )
    }
    soln <- ode.1D( y=yinit, times=times, func=pde.fn, nspec=2 ) # note FIRST COLUMN IS TIME
    attr(soln,"u") <- u
    attr(soln,"v") <- v
    return(soln)
}
cline.from.soln <- function (soln, grid, clinefn=attr(soln,"u"), ... ) {
    A.inds <- 1+(1:grid$N)
    B.inds <- 1+grid$N+(1:grid$N)
    sel.cline <- clinefn(grid$x.mid,s=s)
    pde.cline <- cbind( pde.soln[,1], pde.soln[,A.inds]*sel.cline[col(pde.soln[,A.inds])] + pde.soln[,B.inds]*(1-sel.cline)[col(pde.soln[,B.inds])] )
    attributes(pde.cline) <- c( attributes(pde.cline), attributes(pde.soln)[setdiff(names(attributes(pde.soln)),c("dim","dimnames"))] )
    attr(pde.cline,"nspec")<-1
    return(pde.cline)
}
```

@@

We'll setup, run, and summarize simulations
as described in [this file](../sims/run-sims.html).
```{r setup_sims, cache=TRUE}
source("../sims/sim-fns.R")
zone.age	= 50
ndemes 		= 500
deme.size 	= 10
ninds 		= ndemes*deme.size
SIGMA 		= 10
s           = 0.1
qtl         = list(chr1=data.frame(traits=c("underdominant"), s = s,pos=c(0.5)))
# spatial locations
xx <- (1:ndemes)-0.5-ndemes/2
# run the simulation
sim <- sim.zone(n.gen=zone.age,n.ind=ninds,n.deme=ndemes,sigma=SIGMA,quiet=TRUE)
# post-process
sim.sums <- spBreaks(sim)
```
and, postprocess:
```{r genotype_sims, cache=TRUE, depends="setup_sims"}
rvals <- c(0,exp(seq(log(.001),log(0.5),length.out=10)))
loci <- list(0.5+rvals)
genotypes <- data.frame(do.call(cbind,lapply(sim.sums$ind.ancest, geno.ind, loci)))
freqs <- apply(genotypes,1,function(Z){
            tapply(Z,cut(1:ncol(genotypes),breaks=seq(0,ncol(genotypes),deme.size)),mean)/2
        })
```

@@

```{r match_pde}
tt <- seq(0,zone.age,length.out=100)
pde.solns <- lapply(rvals, function (r) { solve.pde( u=tanh_cline, r=r, times=tt, s=s ) })
```

>>>>

Here's a function to solve this numerically:
```{r solve_pde_setup, cache=TRUE}
source("cline-fns.R")
thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 5000)
solve.pde
cline.from.soln
```

@@

We'll setup, run, and summarize simulations
as described in [this file](../sims/run-sims.html).
```{r setup_sims, cache=TRUE}
source("../sims/sim-fns.R")
zone.age	= 20
ndemes 		= 500
deme.size 	= 5
ninds 		= ndemes*deme.size
SIGMA 		= 10
s           = 0.1
qtl         = list(chr1=data.frame(traits=c("underdominant"), s = s,pos=c(0.5)))
# spatial locations
xx <- (1:ndemes)-0.5-ndemes/2
# run the simulation
sim <- sim.zone(n.gen=zone.age,n.ind=ninds,n.deme=ndemes,sigma=SIGMA,quiet=TRUE)
# post-process
sim.sums <- spBreaks(sim)
```
and, postprocess:
```{r genotype_sims, cache=TRUE, depends="setup_sims"}
rvals <- c(0,exp(seq(log(.001),log(0.5),length.out=10)))
loci <- list(0.5+rvals)
genotypes <- data.frame(do.call(cbind,lapply(sim.sums$ind.ancest, geno.ind, loci)))
freqs <- apply(genotypes,1,function(Z){
            tapply(Z,cut(1:ncol(genotypes),breaks=seq(0,ncol(genotypes),deme.size)),mean)/2
        })
```

@@

Here's the PDE solution:
```{r match_pde}
tt <- seq(0,zone.age,length.out=100)
pde.solns <- lapply(rvals, function (r) { solve.pde( u=function(...)1-tanh_cline(...), r=r, times=tt, s=s, grid=thegrid ) })
pde.clines <- lapply( pde.solns, cline.from.soln, grid=thegrid )
match.clines <- sapply(pde.clines,function(x)x[length(tt),])
```
Now, comparing the PDE solution to the others:
```{r match_cline_pde}
for (k in seq_along(rvals)) {
    plot(xx,freqs[,k],type='l',xlab="spatial location",ylab="proportion",main=sprintf("r=%0.3f",rvals[k]))
    lines(xx,1-tanh_cline(xx,s=s,sigma=SIGMA),col='red',lwd=3)
    lines(1-thegrid$x.mid*SIGMA,match.clines[,k],col='green',lwd=3)
}
```
